### 2009-04-10 doc/storage_system/arc-storage-documentation.tex
added gateway to manual and a small change in the documentation
TODO done
Alex comments up to section 3.4 + Style file for tracking changes

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2CM}
\hypersetup{
  pdfauthor = {Zsombor Nagy},
  pdftitle = {Documentation of the ARC storage system},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

The ARC storage system is a distributed system for storing replicated \emph{file}s on several file storage nodes and manage them in a global namespace.  The files can be grouped into \emph{collection}s (a concept very similar to directories in the common file systems), and a collection can contain sub-collections and sub-sub-collections in any depth. There is a dedicated \emph{root collection} to gather all collections to the global namespace. This hierarchy of collections and files can be referenced using \emph{Logical Name}s (\emph{LN}s). The users can use this global namespace as they were using a local filesystem. Files can be transfered by multiple transfer protocols, and the client side tools hide this from the user. The replicas of the files are stored on different storage nodes. A storage node here is a network-accessible computer having storage space to share, and a storage element service running (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), \url{https://forge.gridforum.org/projects/byteio-wg/}}, etc.). For each storage node one of the services of the ARC storage system is needed to manage it and to integrate it into the system. There is a way on the client side to access third-party storage solutions through the namespace of the ARC storage system. The main services of the storage system are the following (see Figure~\ref{fig:services}):
\begin{itemize}
    \item the \textbf{A-Hash} service, which is a replicated database which is used by the Librarian to store metadata;
    \item the \textbf{Librarian} service, which handles the metadata and hierarchy of collections and files, the location of replicas, and health data of the Shepherd services, using the A-Hash as database;
    \item the \textbf{Bartender} service, which provides a high-level interface for the users and for other services;
    \item the \textbf{Shepherd} service, which manages storage element services, and provides a simple interface for storing files on storage nodes.
\end{itemize}

\begin{figure}[ht]
\centering{{\scalebox{0.9}{\includegraphics{arc-storage-services.pdf}}}
\caption{\label{fig:services}The components of the ARC storage: the \textbf{A-H}ash service, the \textbf{L}ibrarian service, the \textbf{B}artender service and the \textbf{S}hepherd service.}}
\end{figure}

@@

The storage system is capable of storing files which can be grouped in collections and sub-collections, etc.
Every file and collection has a unique ID in the system called the \emph{GUID}. Compared to the well-known structure of local file systems, these GUIDs are very similar to the concept of \emph{inode}s. And as a directory on a local filesystem is basically just a list of name and inode pairs, a collection on the ARC storage is just a list of name and GUID pairs. There is a dedicated collection which is the \emph{root collection}. This makes the namespace of the ARC storage system a hierarchical namespace where you can start at the root collection, and go to sub-collections and sub-sub-collections to get to a file. This path is called the \emph{Logical Name} (LN). For example if there is a sub-collection called \verb!saturn! in the root collection, and there is a file called \verb!rings! in this sub-collection, then the LN of this file is \verb!/saturn/rings!.

Besides the Logical Names we can refer to a file or collection by simply its GUID, or we can use GUIDs and Logical Names together, as seen on Figure~\ref{fig:namespace}.

@@

Example on Figure~\ref{fig:namespace}: if we have a collection with GUID \verb#1234#, and there is a collection called \verb#green# in it, and in \verb#green# there is another collection called \verb#orange#, and in \verb#orange# there is a file called \verb#huge#, then we can refer to this file with the Logical Name \verb#1234/green/orange/huge#, which means that from the collection called \verb!1234! we have to follow along the path: \verb!green!, \verb!orange!, \verb!huge!.

There is a dedicated root collection (which has the GUID \verb#0#), and if a LN starts with no GUID prefix, it is implicitly prefixed with the GUID of this well-known root collection, e.g.~\verb#/why/blue# means \verb#0/why/blue#. If a user wants to find the file called \verb#/why/blue#, the system knows where to start the search: the GUID of the root collection. The root collection knows the GUID of \verb#why#, and the (sub-)collection \verb#why# knows the GUID of \verb#blue#. If the GUID of this file is \verb#5678#, and somebody makes another entry in collection \verb#/why# (= \verb#0/why#) with name \verb#red# and GUID \verb#5678#, then the \verb#/why/red# LN points to the same file as \verb#/why/blue#, which concept is very similar to a hardlink in a regular local file system.

@@

The collections in the ARC storage are logical entities, the content of a collection is stored as metadata of the collection, which means the a collection actually has no physical data. A file however has both metadata and real physical data (the actual bytes of the file). The metadata of a file is stored in the same database where the collections are stored, but the physical data of a file is stored on storage nodes as multiple replicated copies.

A storage node consists of two things: a storage element service which is capable of storing and serving files through a specific protocol (e.g. a web server, an FTP server, a GridFTP server, etc.) and a Shepherd service which provides a simple interface to access the storage node, and which can initiate and manage file transfers through the storage element service. The Shepherd has different backends for the supported storage element services which made it possible the communicate with them.

So we have logical files, which are part of the hierarchical namespace and have a GUID and other metadata, and a logical file has one or more physical replicas. The physical replicas stored on separate storage nodes. In order to connect the logical file to its replicas, we need to have some pointers. Each storage node has a URL and each replica has a unique ID within the storage node called \emph{referenceID}, the URL and the referenceID together is called a \emph{Location}, a Location unambiguously points to one specific replica. So to connect the logical files to the physical ones, each logical file has a list of Locations.

The user can specify for each file how many replicas are needed. The ARC storage system periodically checks the number of replicas, and automatically creates new replicas if there are fewer then needed, and removes replicas if there are more.

The different replicas of a file could be in different states, e.g. the replica could be valid and alive or just in the process of creation or it could be corrupt or a whole storage node could be offline. This state is always stored as metadata next to the Location of the given replica. For each file there is a checksum calculated, and this checksum is used to detect if a replica gets corrupted. If a storage node (more precisely: the Shepherd service on the storage node) detects that a file is invalid, it reports this so the metadata will be in sync with the real state. And the storage nodes sends heartbeat messages periodically, and if a storage node goes offline, the missing heartbeat triggers the modification of metadata as well.

@@

The A-Hash is the replicated\footnote{Depends on system configuration. It is possible to configure the storage system with only one central A-Hash too} metadata store of the storage system. It consistently stores `objects' which contain property-value pairs organized in sections. All metadata about files and collections in the storage system are stored in the A-Hash. Additionall system configutation information (e.g. about A-Hash replication, about Shepherd services, etc.) is stored in it as well. The A-Hash itself does not interpret any data.

@@

The Librarian manages the storage system hierarchy of files and collections including their metadata. It can traverse Logical Names and return the corresponding metadata. The Shephards send their heartbeat messages to the Librarians which if needed respond by modifying the states of files stored on a certain storage node. The Librarian itself is a stateless service, it uses the A-Hash to store and retrieve the metadata and thus there can be any number of independent Librarian services (all using the same A-Hashes). This ensures high-availability and load-balancing of the storage system.

@@

The Bartender service provides the high-level interface of the storage system to the clients. Every interaction between a client and the ARC storage system starts by sending a request to a Bartender. Using the Bartender it is possible to create and remove collections, create, get and remove files, move files and collections within the namespace using Logical Names, etc. The Bartender authorizes users and apply access policies of files and collections. It communicates with the Librarian and Shepherd services to accomplish the client’s requests. The file data itself does however not go through the Bartender; file transfers are directly performed between the storage nodes and the clients. There could be any number of independent Bartender services in the system which ensure high-availability and load-balancing. The Bartender also provides a way to access files on third-party storage solutions through its interface by mounting the namespace of the third-party storage into the namespace of the ARC storage (this is accomplished by so called `gateway' modules).

@@

The Shepherd service manages the storage element service on the storage node. It reports its health state to a Librarian and provides the interface for initiating file transfers. For each kind of storage element service (e.g. a HTTP server, an FTP server, a storage solution with a GridFTP interface, etc.) a Shepherd backend is needed which is capable of managing the given storage element service. The Shepherd service periodically checks the health of the replicas using their checksums, and if a replica is deleted or corrupted, the Shepherd tries to recover it by downloading a valid copy from another storage node. The Shepherd also periodicallys check if a file has the correct number of replicas in the system. If too few or too many replicas are found, new replicas will be created or redundant replicas removed.

@@

The ARC storage system consists of several services (both different kinds of services and multiple instances of the same kind). The A-Hash, Librarian and Shepherd services are `internal' services in the sense that the end-user of the storage system never communicates with them directly. But these internal services are communicating with each other, thus they have to know which other services to trust. We call this aspect of the security architecture of the ARC storage `inter-service authorization'.

The end users always connect to one of the Bartender services, which will decide which permission the user has. This is the `high-level authorization' part of the security architecture.

For transferring the actual file data, the users have to connect to storage element services running on storage nodes. These services also have their own authentication and authorization methods. Managing these aspects is the `transfer-level authorization' part of the security architecture of the ARC storage.

@@

In a deployment of the ARC storage system it is possible to have several A-Hash, Librarian, Shepherd and Bartender services. The Bartenders send requests to the Librarians and the Shepherds, the Shepherds communicate with the Librarians, the Librarians talk with the A-Hashes. If any of these services gets compromised or a new rouge service is inserted in the system, security is lost. That's why it is vital for each service to authorize other services before sending or accepting requests. The services communicate via the HTTPS protocol, which means that they should provide an X.509 certificate for each connection, and that they can examine eachother's certificates. Because of these X.509 certificates each service has Distinguish Name (DN), which can be used for identifying trusted services. A list of trusted DNs can be configured into each service, or a list of trusted services can be stored on a remote location. The services will only accept connections if the DN of the other end is listed in this list of trusted DNs. However the Bartender services will accept any incoming connection, which are from the users, because the users are authenticated differently.

@@

The Librarian component of the ARC storage system stores all the metadata about files and collections. For each file and collection there are access policies in the form of access control rules, and these are stored among the metadata. The users are identified by their DNs, and an access control rule specify the rights of the given user. A rule can be represented like this:

@@

Besides specifying only one user with a DN, there are other types of access control rules: we can have a rule for a whole VO (Virtual Organization) or for ALL users, like this:

\begin{verbatim}
    ALL +action
    VOMS:knowarc.eu +action -action 
\end{verbatim}

@@

Currently the transfer-level authorization is kept very simple. When the Bartender decides that a user has permission to download a file, then the Bartender chooses a replica, and initiates the transfer. The result of this initiation is a URL which is called the transfer URL (TURL). This TURL is unique for each request, even for request to the same replica, and this TURL is only valid for one download. Currently we configure the storage element services to not do any authorization, and we use these one-time URLs to ensure that only the authorized users can access the contents of the storage elements.

@@

\chapter{Use cases} % (fold)
\label{cha:use_cases}

@@

We want to list the contents of a collection, which has a Logical Name of \verb!/ourvo/users/we!. In order to do this, we have to contact a Bartender, and send a request to it containing the Logical Name we want to list, and the response from the Bartender will contain the list of entries. The steps are represented on Figure~\ref{fig:listing}.

\begin{enumerate}
    \item We need to know the URL of a Bartender. This could be preconfigured on the client side or in future releases it could be acquired from an information system. When we have the URL, we send a `list' request which contains the Logical Name \verb!/ourvo/users/we!.
    \item The Bartender tries to get the metadata of the given LN by sending a `traverseLN' request to a Librarian.
    \item The Librarian service starts the traversing by asking an A-Hash service about the first part of the LN, which is the \verb!/! root collection. The A-Hash service only knows about GUIDs and not about LNs, but the GUID of the root collection is well-known, so the A-Hash can return the metadata of it which contains the list of files and sub-collections in the root collection.
    \item Hopefully the \verb!ourvo! collection can be found in the root collection, which means that new the Librarian knows its GUID, and can ask for its metadata from the A-Hash.
    \item After the A-Hash returns the metadata of the \verb!/ourvo! collection, the Librarian finds the GUID of \verb!users! in it, then gets its metadata.
    \item The A-Hash returns the metadata of \verb!/ourvo/users! which contains the GUID of \verb!we!, so the Librarian can ask for its metadata.
    \item At last the A-Hash returns the metadata of \verb!/ourvo/users/we! to the Librarian, and the Librarian returns it to the Bartender. This metadata contains the list of entries within this collection, and it also contains the access policies for this collection.
    \item The Bartender first checks if based on our DN (or our VO membership) and the access policies of this collection do we have rights to get the contents of this collection or not. If we are approved, then because the `list' request should return additional metadata about each entry in the collection, the Bartender send a `get' message to a Librarian requesting metadata of all the entries in this collection.
    \item The Librarian gets the data from the A-Hash and returns it to the Bartender.
    \item Now the Bartender has all the needed information, so it could return the proper response to us, the user. Our client tool formats and prints the results nicely.
\end{enumerate}

@@

In this use case we want to download a file which has a Logical Name of \verb#/ourvo/users/we/thefilewewant# (see Figure~\ref{fig:downloading}).

\begin{enumerate}
    \item We connect a Bartender and send a `getFile' request with the LN of our file.
    \item The Bartender contacts a Librarian to traverse the Logical Name and to get the metadata of our file.
    \item The Librarian do the step by step traversing and gets all the data from the A-Hash and returns the metadata of our file to the Bartender. This metadata contains the location of the file's replicas and the access policies of this file.
    \item The Bartender checks based on the access policies and our identity if we are allowed to get the file, and if we are good to go, then it chooses a replica location. A location consists of the URL of a Shepherd service, and the ID of the replica within that Shepherd (which is called a `referenceID'). The Bartender sends a `get' request to the chosen Shepherd.
    \item The Shepherd prepares the file transfer by asking the storage element service to create a new one-time URL for this replica. This URL called the transfer URL, and it will only be valid for one download. The Shepherd returns the TURL to the Bartender.
    \item The Bartender returns the TURL to us.
    \item Now we use this TURL to get the file directly from the storage element service on the storage node.
\end{enumerate}

@@

We want to create a new (empty) collection as sub-collection of \verb#/ourvo/common#, and we want to call it \verb!docs! (see Figure~\ref{fig:makecollection})

\begin{enumerate}
    \item We contact a Bartender and send a `makeCollection' request with the LN \verb!/ourvo/common/docs!.
    \item The Bartender asks a Librarian to traverse this LN.
    \item The Librarian try to traverse the Logical Name and it stops at the last possible point and returns the metadata of the last element. Because we want to put our collection to a new path but into an existing collection, we expect that only the \verb#/ourvo/common# part of the LN can be traversed. If the Librarian could traverse the whole LN that would mean that there is already an existing file or collection by that name. If the parent collection does exist then the Librarian can traverse it which means that the parent collection's metadata is returned to the Bartender. 
    \item The Bartender checks the access policies to decide if we have permissions to put something into this collection. Then it asks the Librarian to create a new collection.
    \item The Librarian creates the collection, and returns its GUID. At this point this new collection has no real Logical Name yet, it only has a GUID, but it is not yet put into its parent collection.
    \item The Bartender now asks the Librarian to add this new entry into the parent collection, which means that the new GUID and the name \verb!docs! are added as a pair.
    \item The Librarian returns with a status message.
    \item Finally the Bartender tells us if everything went OK or not.
\end{enumerate}

@@

\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc-storage-removing.pdf}}}
\caption{\label{fig:removing}Removing a file} }
\end{figure}
\begin{enumerate}
    \item If we want to remove a file, we should connect to a Bartender with the LN of the file we want to remove.
    \item The Bartender asks the Librarian to traverse the LN.
    \item The Librarian returns the metadata of the file. The metadata contains information about all the hardlinks which points to this file if there are more than one.
    \item Now the Bartender asks the Librarian to remove the file.
    \item After that, the Bartender asks the Librarian to remove the links to this file from all the parent collections.
    \item Next time a Shepherd which has a replica of this file does its periodic check, it asks the Librarian about the file, and notices that the file does not exist anymore, so it removes the replica itself from the storage node.
\end{enumerate}

@@

The ARC storage system stores different kinds of metadata about files, collections, mount point, Shepherd services, etc. Each of these has a unique ID which we call `GUID'.

>>>>

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
\usepackage{trackchanges}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2CM}
\hypersetup{
  pdfauthor = {Zsombor Nagy},
  pdftitle = {Documentation of the ARC storage system},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

Chelonia is a distributed system for storing replicated \emph{file}s on several file storage nodes and managing them in a global namespace.  The files can be grouped into \emph{collection}s (a concept very similar to directories in the common file systems). A collection can contain sub-collections and sub-sub-collections in any depth. There is a dedicated \emph{root collection} to gather all collections to the global namespace. This hierarchy of collections and files can be referenced using \emph{Logical Name}s (\emph{LN}'s). The users can use this global namespace as if they were using a local filesystem. Files can be transfered by several different transfer protocols. The client side tools hide this from the user. The replicas of the files are stored on different storage nodes. A storage node here is a network-accessible computer having storage space to share, and a running storage element service (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), \url{https://forge.gridforum.org/projects/byteio-wg/}}, etc.). \note[JKN]{TODO: Rephrase this sentence; "one of the services" seems vague}For each storage node one of the services of the ARC storage system is needed to manage it and to integrate it into the system. The client side provides access to third-party storage solutions through the namespace of the ARC storage system. The main services of the storage system are the following (see Figure~\ref{fig:services}):
\begin{itemize}
    \item the \textbf{A-Hash} service, which is a replicated database which is used by the Librarian to store metadata;
    \item the \textbf{Librarian} service, which handles the metadata and hierarchy of collections and files, the location of replicas, and health data of the Shepherd services, using the A-Hash as database;
    \item the \textbf{Bartender} service, which provides a high-level interface for the users and for other services;
    \item the \textbf{Shepherd} service, which manages storage element services, and provides a simple interface for storing files on storage nodes.
\end{itemize}

\begin{figure}[ht]
\centering{{\scalebox{0.8}{\includegraphics{arc-storage-services.pdf}}}
\caption{\label{fig:services}{\bf TODO:} explain figure, since components are not yet introduced. The components of the ARC storage: the \textbf{A-H}ash service, the \textbf{L}ibrarian service, the \textbf{B}artender service and the \textbf{S}hepherd service.}}
\end{figure}

@@

The storage system is capable of storing files which can be grouped in collections and sub-collections, etc.
Every file and collection has a unique ID in the system called the \emph{GUID}. Compared to the well-known structure of local file systems, these GUID's are very similar to the concept of \emph{inode}s. And as a directory on a local filesystem is basically just a list of name and inode pairs, a collection on the ARC storage is just a list of name and GUID pairs. There is a dedicated collection which is the \emph{root collection}. This makes the namespace of the ARC storage system a hierarchical namespace where you can start at the root collection, and go to sub-collections and sub-sub-collections to get to a file. This path is called the \emph{Logical Name} (LN). For example if there is a sub-collection called \verb!saturn! in the root collection, and there is a file called \verb!rings! in this sub-collection, then the LN of this file is \verb!/saturn/rings!.

In addition to the Logical Names we can refer to a file or collection simply by its GUID, or we can use GUID's and Logical Names together, as seen in Figure~\ref{fig:namespace}.

@@

An example of the hierarchy of the global namespace is shown in Figure~\ref{fig:namespace}: if we have a collection with GUID \verb#1234#, and there is a collection called \verb#green# in it, and in \verb#green# there is another collection called \verb#orange#, and in \verb#orange# there is a file called \verb#huge#, then we can refer to this file with the Logical Name \verb#1234/green/orange/huge#, which means that from the collection called \verb!1234! we have to follow along the path: \verb!green!, \verb!orange!, \verb!huge!.

There is a dedicated root collection (which has the GUID \verb#0#), and if a LN starts without a GUID prefix, it is implicitly prefixed with the GUID of the root collection, e.g.~\verb#/why/blue# means \verb#0/why/blue#. If a user wants to find the file called \verb#/why/blue#, the system knows where to start the search: the GUID of the root collection. The root collection knows the GUID of \verb#why#, and the (sub-)collection \verb#why# knows the GUID of \verb#blue#. If the GUID of this file is \verb#5678#, and somebody makes another entry in collection \verb#/why# (= \verb#0/why#) with name \verb#red# and GUID \verb#5678#, then the \verb#/why/red# LN points to the same file as \verb#/why/blue#, a concept very similar to a hardlink in a \note[JKN]{Todo: Posix file system?}regular local file system.

@@

The collections in Chelonia are logical entities. The content of a collection is stored as metadata of the collection, which means that a collection actually has no physical data. A file, however, has both metadata and real physical data (the actual content of the file). The metadata of a file is stored in the same database where the collections are stored, but the physical data of a file is stored on storage nodes as multiple replicated copies.

A storage node consists of two things: a storage element service which is capable of storing and serving files through a specific protocol (e.g. a web server, an FTP server, a GridFTP server, etc.) and a Shepherd service which provides a simple interface to access the storage node, and which can initiate and manage file transfers through the storage element service. The Shepherd has different backends for the supported storage element services which made it possible to communicate with them.

A logical file, which are part of the hierarchical namespace, has a GUID and other metadata, and one or more physical replicas. The physical replicas are stored on separate storage nodes. In order to connect the logical file to its replicas, the system needs to have global pointers. Each storage node has a URL and each replica has a unique ID within the storage node called \emph{referenceID}, the URL and the referenceID together is called a \emph{Location}, a Location unambiguously points to one specific replica. To connect the logical files to the physical ones, each logical file has a list of Locations.

The user can specify how many replicas are required for each file. The ARC storage system periodically checks the number of replicas\note[JKN]{TODO: How will this scale?}, and automatically creates new replicas if there are fewer than required, and removes replicas if there are more.

The different replicas of a file could be in different \note[JKN]{TODO: which states? a reference to definition of states maybe?}states, e.g. the replica could be valid and alive or just in the process of creation or it could be corrupt or a whole storage node could be offline. This state is always stored as metadata of the given replica. For each file there is a checksum calculated, and this checksum is used to detect if a replica gets corrupted. If a storage node (more precisely: the Shepherd service on the storage node) detects that a file is invalid, it reports this so the metadata will be in sync with the real state. And the storage nodes sends heartbeat messages periodically, and if a storage node goes offline, the missing heartbeat triggers the modification of metadata as well.

@@

The A-Hash is the metadata store of the storage system. The A-Hash may be deployed as either a centralized or a replicated service. It consistently stores `objects' which contain property-value pairs organized in sections. All metadata about files and collections in the storage system are stored in the A-Hash. Additionall system configutation information (e.g. about A-Hash replication, about Shepherd services, etc.) is stored in it as well. The A-Hash itself does not interpret any data.

@@

The Librarian manages the storage system hierarchy of files and collections including their metadata. It can traverse Logical Names and return the corresponding metadata. The Shepherds (see Sect. \ref{sec:the_shepherds}) send their heartbeat messages to the Librarians and, if needed, the Librarian responds by modifying the states of files stored on a certain storage node. The Librarian itself is a stateless service, it uses the A-Hash to store and retrieve the metadata and thus there can be any number of independent Librarian services (all using the same A-Hashes). This ensures high-availability and load-balancing of the storage system.

@@

The Bartender service provides the high-level interface of the storage system to clients. Every interaction between a client and the ARC storage system starts by sending a request to a Bartender. Using the Bartender it is possible to create and remove collections, create, get and remove files, move files and collections within the namespace using Logical Names, etc. The Bartender authorizes users and applies access policies of files and collections. It communicates with the Librarian and Shepherd services to execute the requests of the clients. The file data itself does not go through the Bartender; file transfers are performed directly between the storage nodes and the clients. Any number of independent Bartender services may be deployed in the system to ensure high-availability and load-balancing. The Bartender also provides access to files in third-party storage solutions through its interface by mounting the namespace of the third-party storage into the namespace of the ARC storage (this is accomplished by so-called `gateway' modules).

@@

The Shepherd service manages the storage element service on the storage node. It reports its health state to a Librarian and provides the interface for initiating file transfers. For each kind of storage element service (e.g. an HTTP server, an FTP server, a storage solution with a GridFTP interface, etc.) a Shepherd backend is needed which is capable of managing the given storage element service. The Shepherd service periodically checks the health of the replicas using their checksums, and if a replica is deleted or corrupted, the Shepherd tries to recover it by downloading a valid copy from another storage node. The Shepherd also periodically checks if a file has the correct number of replicas in the system. If too few or too many replicas are found, new replicas will be created or redundant replicas removed.

@@

Chelonia consists of several services (both different kinds of services and multiple instances of the same kind). The A-Hash, Librarian and Shepherd services are `internal' services in the sense that the end-user of the storage system never communicates with them directly. But these internal services are communicating with each other, thus they have to know which other services to trust. We call this aspect of the security architecture of the ARC storage system `inter-service authorization'.

The end users always connect to one of the Bartender services, which will decide which permissions the user has. This is the `high-level authorization' part of the security architecture.

For transferring the actual file data, the users have to connect to storage element services running on storage nodes. These services have their own authentication and authorization methods. Managing these aspects is the `transfer-level authorization' part of the security architecture of Chelonia.

@@

In a deployment of Chelonia, it is possible to have several A-Hash, Librarian, Shepherd and Bartender services. The Bartenders send requests to the Librarians and the Shepherds, the Shepherds communicate with the Librarians and the Librarians with the A-Hashes. If any of these services gets compromised or a new rogue service is inserted in the system, security is lost. That's why it is vital for each service to authorize other services before sending or accepting requests. The services communicate via the HTTPS protocol, which means that they should provide an X.509 certificate for each connection, and that they can examine each other's certificates. Each service is required to have an X.509 certificate and is thus identified by its Distinguish Name (DN). A list of trusted DN's can be configured into each service, or a list of trusted services can be stored on a remote location. The services will only accept connections if the DN of the other end is listed in this list of trusted DN's. However, the Bartender services will accept any incoming connection, which are from the users, because the users are authenticated differently.

@@

The Librarian component of the ARC storage system stores all the metadata about files and collections. For each file and collection there are access policies in the form of access control rules, and these are stored among the metadata. The users are identified by their DN's, and an access control rule specifies the rights of the given user. A rule can be represented like this:

@@

Besides specifying only one user with a DN, there are other types of access control rules: we can have a rule for a whole VO (Virtual Organization) or for all users, like this:

\begin{verbatim}
    VOMS:knowarc.eu +action -action 
    ALL +action
\end{verbatim}

@@

Currently the transfer-level authorization is very simple. When the Bartender decides that a user has permission to download a file, then the Bartender chooses a replica, and initiates the transfer. The result of this initiation is the transfer URL (TURL). This TURL is unique for each request, even for requests for the same replica, and this TURL is only valid for one download. Currently we configure the storage element services to not do any authorization, and we use these one-time TURL's to ensure that only authorized users can access the contents of the storage elements.

@@

\chapter{Implementation of use cases} % (fold)
\label{cha:use_cases}

\note[JKN]{TODO: Short intro to this chapter mentioning that this chapter contains a partial list of use cases and that not all features are covered.}

@@

A user wants to list the contents of a collection, which has a Logical Name of \verb!/ourvo/users/user!. In order to do this, the client tool has to contact a Bartender, and send a request to it containing the Logical Name the user wants to list, and the response from the Bartender will contain the list of entries. The steps are represented in Figure~\ref{fig:listing}.

\begin{enumerate}
    \item The client tool needs to know the URL of a Bartender. This could be preconfigured on the client side or in future releases it could be acquired from an information system. When the client tool has the URL, it sends a `list' request which contains the Logical Name \verb!/ourvo/users/user!.
    \item The Bartender tries to get the metadata of the given LN by sending a `traverseLN' request to a Librarian.
    \item The Librarian service starts the traversing by asking an A-Hash service about the first part of the LN, which is the \verb!/! root collection. The A-Hash service only knows about GUID's and not about LN's, but the GUID of the root collection is well-known, so the A-Hash can return the metadata of the list of files and sub-collections in the root collection.
    \item Hopefully, the \verb!ourvo! collection can be found in the root collection, which means that the Librarian knows its GUID, and can ask for its metadata from the A-Hash.
    \item After the A-Hash returns the metadata of the \verb!/ourvo! collection, the Librarian finds the GUID of \verb!users! in it, then gets its metadata.
    \item The A-Hash returns the metadata of \verb!/ourvo/users! which contains the GUID of \verb!user!, so the Librarian can ask for its metadata.
    \item At last the A-Hash returns the metadata of \verb!/ourvo/users/user! to the Librarian, and the Librarian returns it to the Bartender. This metadata contains the list of entries within this collection, and it also contains the access policies for this collection.
    \item The Bartender first checks, based on the user's DN (or his VO membership) and the access policies of this collection, if the user has permission to get the contents of this collection or not. If the user is approved, then because the `list' request should return additional metadata about each entry in the collection, the Bartender sends a `get' message to a Librarian which requests the metadata of all the entries in this collection.
    \item The Librarian gets the metadata from the A-Hash and returns it to the Bartender.
    \item Now the Bartender has all the needed information and returns the proper response to the client tool which formats and prints the results nicely for the user.
\end{enumerate}

@@

In this use case a user wants to download a file which has a Logical Name of \verb#/ourvo/users/user/thefilehewants# (see Figure~\ref{fig:downloading}).

\begin{enumerate}
    \item The client tool connects to a Bartender and sends a `getFile' request with the LN of the file.
    \item The Bartender contacts a Librarian to traverse the Logical Name and to get the metadata of our file.
    \item The Librarian does the step-by-step traversing of the name space and gets all the metadata from the A-Hash and returns the metadata of our file to the Bartender. This metadata contains the location of the file's replicas and the access policies of this file.
    \item The Bartender checks, based on the access policies and the user's identity, if he is allowed to get the file, and if allowed, it chooses \note[JKN]{TODO: We need to define the choosing algorithm somewhere} a replica location. A location consists of the URL of a Shepherd service, and the ID of the replica within that Shepherd (which is called a `referenceID'). The Bartender sends a `get' request to the chosen Shepherd.
    \item The Shepherd prepares the file transfer by asking the storage element service to create a new one-time TURL for this replica. The Shepherd returns the TURL to the Bartender.
    \item The Bartender returns the TURL to the client tool.
    \item The client tool use this TURL to get the file directly from the storage element service on the storage node.
\end{enumerate}

@@

Here, the user wants to create a new (empty) collection as sub-collection of \verb#/ourvo/common#, and he wants to call it \verb!docs! (see Figure~\ref{fig:makecollection})

\begin{enumerate}
    \item The client tool contacts a Bartender and send a `makeCollection' request with the LN \verb!/ourvo/common/docs!.
    \item The Bartender asks a Librarian to traverse this LN.
    \item The Librarian tries to traverse the Logical Name and it stops at the last possible point and returns the metadata of the last element. Because the user wants to put his collection in a new path but into an existing collection, we expect that only the \verb#/ourvo/common# part of the LN can be traversed. If the Librarian could traverse the whole LN that would mean that there is already an existing file or collection by that name. If LN does not exist but the the parent collection does, then the parent collection's metadata is returned to the Bartender. 
    \item The Bartender checks the access policies to decide if the user has permissions to put something into this collection. Then it asks the Librarian to create a new collection.
    \item The Librarian creates the collection, and returns its GUID. At this point this new collection has no real Logical Name, it only has a GUID, but it is not yet put into its parent collection.
    \item The Bartender now asks the Librarian to add this new entry into the parent collection, which means that the new GUID and the name \verb!docs! are added as a pair.
    \item The Librarian returns with a status message.
    \item Finally the Bartender tells the client tool if everything went OK or not.
\end{enumerate}

@@

\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc-storage-removing.pdf}}}
\caption{\label{fig:removing}Removing a file} }
\end{figure}
In Chelonia, files are deleted through so called {\it lazy deletion}, meaning that at first only the Logcial Name (LN) is deleted, while the physical replicas are removed at a later stage.
\begin{enumerate}
    \item If the user wants to remove a file, his client tool should connect to a Bartender with the LN of the file he wants to remove.
    \item The Bartender asks the Librarian to traverse the LN.
    \item The Librarian returns the metadata of the file. The metadata contains information about all the hardlinks which points to this file if there are more than one.
    \item Now the Bartender asks the Librarian to remove the file.
    \item After that, the Bartender asks the Librarian to remove the links to this file from all the parent collections.
    \item The next time a Shepherd which has a replica of this file does its periodic check, it asks the Librarian about the file, and notices that the file does not exist anymore, so it removes the replica itself from the storage node.
\end{enumerate}

@@

The ARC storage system stores different kinds of metadata about files, collections, mount points, Shepherd services, etc. Each of these has a unique ID; a GUID.
